Are environmental and geographic effective surrogates for genetic variation in conservation planning?

### 2016-01-01 article/article.Rmd
- set parameters to TOML format - make parameters loaded only in when needed to reduce overheads if parameters are changed

First, we grouped individuals into genetic lineages for each species (Figure S2). The purpose of this step was to avoid confounding variation between populations with due to genetic history with variation attributed to selection pressures. We used the Bayesian clustering method implemented in $\texttt{Structure}$ [version 2.3.4; @r461; @r462] assuming admixture and an independent alleles model. For each species, we identified the most plausible number of populations (K = `r st.k[1]`--`r st.k[2]`; `r st.numruns` runs per K; `r st.burnin` burn-in iterations; `r st.numreps` post-burn-in iterations) using the $\Delta$K method [Figure S3; @r460]. We aggregated the population membership probabilities from replicate runs using $\texttt{CLUMPP}$ [version 1.1.2; large K greedy search method with `r cl.repeats` random starts per K per species; Figure S4; @r459]. Individuals with a moderate probability of belonging to a single population ($\geq `r st.probthresh`$) were assigned accordingly (Figure S5), and subject to subsequent outlier locus detection methods.

We used $\texttt{Bayescan}$ [version 2.1; @r456] to identify loci under selection for each species (Table S2). For each species, $\texttt{Bayescan}$ was run using the population memberships identified in the previous analysis (1:1 prior odds; `r bs.nbp` pilot runs; `r bs.burn` burn-in iterations; `r bs.n` post-burn-in iterations thinned by `r bs.thin` iterations). Following guidelines in the $\texttt{Bayescan}$ manual, we omitted loci where the global frequency of the minor allele was $\geq 0.05$ to avoid false-positives. To ensure convergence, we ran `r bs.reps` replicates per species using a suitable false discovery rate [FDR; $\leq `r bs.fdr`$; @r466].

After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if present) and neutral genetic variation for each species (Figures S4--S29). Note that while individuals not assigned to populations were omitted from the $\texttt{Bayescan}$ analysis, here we used all the individuals. For each species, for each type of genetic variation present, we used Gower distances [@r464] to express the genomic differences between individuals [using the cluster R package to accommodate sparsity; @r457]. These distances were subject to a non metric multi-dimensional scaling analysis [implemented in the vegan R package using K=`r mds.k` and `r mds.trymax` random starts; Table S3; @r458]. The average of the ordinated values associated with individuals in each grid cell we used to characterize the genomic variation in the grid cell.

### Prioritizations
We used the unreliable representative and adequate prioritization (URAP) formulation of the reserve selection problem [@r465]. We treated the genomic and surrogate variables as distinct attribute spaces. We set the demand points for each species and its attribute space using the values associated with the planning units. First, we generated a set of single species prioritizations. For each species, we generated a prioritization using a `r (rapr.amount.target*100)`% amount-based target, a prioritization using a `r (rapr.amount.target*100)`% amount-based target and `r (rapr.surrogate.target*100)`% space-based targets for the surrogate variation, and a prioritization using a `r (rapr.amount.target*100)`% amount-based target and `r (rapr.genetic.target*100) `% space-based targets for the genomic variation. Next, we generated three multi-species prioritizations using the same sets of targets used in the single species prioritizations. Finally, to understand the relationship between the surrogate-based and genomic-based targets, we generated a suite of prioritizations using increasing space-based targets between `r (min(rapr.pareto.surrogate.targets)*100)`% and `r (max(rapr.pareto.surrogate.targets)*100)`% in increments of `r diff(rapr.pareto.surrogate.targets[1:2])*100`%. For each prioritization we calculated the proportion of the adaptive (if present) and neutral genetic variation that it secured. We solved the reserve selection problems to within `r (gb.MIPGap*100)`% of optimality using the $\texttt{rapr}$ R package [@r465] and $\texttt{Gurobi}$ [@r463].

>>>>

First, we grouped individuals into genetic lineages for each species (Figure S2). The purpose of this step was to avoid confounding variation between populations with due to genetic history with variation attributed to selection pressures. We used the Bayesian clustering method implemented in $\texttt{Structure}$ [version 2.3.4; @r461; @r462] assuming admixture and an independent alleles model. For each species, we identified the most plausible number of populations (K = `r min(structure.params.LST[[MODE]]$k)`--`r max(structure.params.LST[[MODE]]$k)`; `r structure.params.LST[[MODE]]$numruns` runs per K; `r structure.params.LST[[MODE]]$burnin` burn-in iterations; `r structure.params.LST[[MODE]]$numreps` post-burn-in iterations) using the $\Delta$K method [Figure S3; @r460]. We aggregated the population membership probabilities from replicate runs using $\texttt{CLUMPP}$ [version 1.1.2; large K greedy search method with `r clumpp.params.LST[[MODE]]$repeats` random starts per K per species; Figure S4; @r459]. Individuals with a moderate probability of belonging to a single population ($\geq `r structure.params.LST[[MODE]]$probthresh`$) were assigned accordingly (Figure S5), and subject to subsequent outlier locus detection methods.

We used $\texttt{Bayescan}$ [version 2.1; @r456] to identify loci under selection for each species (Table S2). For each species, $\texttt{Bayescan}$ was run using the population memberships identified in the previous analysis (1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations). Following guidelines in the $\texttt{Bayescan}$ manual, we omitted loci where the global frequency of the minor allele was $\geq 0.05$ to avoid false-positives. To ensure convergence, we ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate [FDR; $\leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].

After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if present) and neutral genetic variation for each species (Figures S4--S29). Note that while individuals not assigned to populations were omitted from the $\texttt{Bayescan}$ analysis, here we used all the individuals. For each species, for each type of genetic variation present, we used Gower distances [@r464] to express the genomic differences between individuals [using the cluster R package to accommodate sparsity; @r457]. These distances were subject to a non metric multi-dimensional scaling analysis [implemented in the vegan R package using K=`r mds.params.LST[[MODE]]$k` and `r mds.params.LST[[MODE]]$trymax` random starts; Table S3; @r458]. The average of the ordinated values associated with individuals in each grid cell we used to characterize the genomic variation in the grid cell.

### Prioritizations
We used the unreliable representative and adequate prioritization (URAP) formulation of the reserve selection problem [@r465]. We treated the genomic and surrogate variables as distinct attribute spaces. We set the demand points for each species and its attribute space using the values associated with the planning units. First, we generated a set of single species prioritizations. For each species, we generated a prioritization using a `r (rapr.params.LST[[MODE]]$amount.target*100)`% amount-based target, a prioritization using a `r (rapr.params.LST[[MODE]]$amount.target*100)`% amount-based target and `r (rapr.params.LST[[MODE]]$surrogate.target*100)`% space-based targets for the surrogate variation, and a prioritization using a `r (rapr.params.LST[[MODE]]$amount.target*100)`% amount-based target and `r (rapr.params.LST[[MODE]]$genetic.target*100) `% space-based targets for the genomic variation. Next, we generated three multi-species prioritizations using the same sets of targets used in the single species prioritizations. Finally, to understand the relationship between the surrogate-based and genomic-based targets, we generated a suite of prioritizations using increasing space-based targets between `r (min(rapr.params.LST[[MODE]]$pareto.surrogate.targets)*100)`% and `r (max(rapr.params.LST[[MODE]]$pareto.surrogate.targets)*100)`% in increments of `r diff(rapr.params.LST[[MODE]]$pareto.surrogate.targets[1:2])*100`%. For each prioritization we calculated the proportion of the adaptive (if present) and neutral genetic variation that it secured. We solved the reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)`% of optimality using the $\texttt{rapr}$ R package [@r465] and $\texttt{Gurobi}$ [@r463].

### 2016-01-01 article/article.tex
- set parameters to TOML format - make parameters loaded only in when needed to reduce overheads if parameters are changed

\setlength{\droptitle}{-2em}
  \title{Are environmental and geographic effective surrogates for genetic
variation in conservation planning?}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{Jeffrey O. Hanson$^1$, Jonathan R. Rhodes$^2$, Cynthia Riginos$^1$, Hugh
P. Possingham$^1$, Richard A. Fuller$^1$\\$^1$School of Biological
Sciences, The University of Queensland, Brisbane, QLD,
Australia\\$^2$School of Geography, Planning and Environmental
Management, The University of Queensland, Brisbane, QLD,
Australia\\Correspondance should be addressed to
\href{mailto:jeffrey.hanson@uqconnect.edu.au}{jeffrey.hanson@uqconnect.edu.au}}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{24 December 2015}

% load packages
\usepackage{amsmath,amsfonts,float,makecell,titletoc,tocloft,titlesec,natbib}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[utf8]{inputenc}

@@

First, we grouped individuals into genetic lineages for each species
(Figure S2). The purpose of this step was to avoid confounding variation
between populations with due to genetic history with variation
attributed to selection pressures. We used the Bayesian clustering
method implemented in $\texttt{Structure}$ (version 2.3.4; Pritchard
\emph{et al.} 2000; Falush \emph{et al.} 2007) assuming admixture and an
independent alleles model. For each species, we identified the most
plausible number of populations (K = 1--2; 4 runs per K; 10 burn-in
iterations; 10 post-burn-in iterations) using the $\Delta$K method
(Figure S3; Evanno \emph{et al.} 2005). We aggregated the population
membership probabilities from replicate runs using $\texttt{CLUMPP}$
(version 1.1.2; large K greedy search method with 1000 random starts per
K per species; Figure S4; Jakobsson and Rosenberg 2007). Individuals
with a moderate probability of belonging to a single population
($\geq 0.75$) were assigned accordingly (Figure S5), and subject to
subsequent outlier locus detection methods.

@@

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

\setlength{\droptitle}{-2em}
  \title{}
  \pretitle{\vspace{\droptitle}}
  \posttitle{}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\subsection{Figures}\label{figures}

\begin{figure}[htbp]
\centering
\includegraphics{figures_files/figure-latex/unnamed-chunk-2-1.pdf}
\caption{Summary of single species prioritisations. Single species
prioritisations were generated using amount-based targets, amount-based
and surrogate-based targets, and amount-based and genetic-based targets
for each species. Data shows the performance of prioritisations
generated using these three sets of targets. Bars denote means and
standard errors.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figures_files/figure-latex/unnamed-chunk-3-1.pdf}
\caption{Multi-species prioritisations. Panel (a) shows the
prioritisation generated for using just amount-based targets. Panel (b)
shows the prioritisation generated using amount-based and surrogate
based targets. Panel (c) shows the prioritisation generated using
amount-based and genetic-based targets}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figures_files/figure-latex/unnamed-chunk-4-1.pdf}
\caption{Summary of multi-species prioritisations. Three prioritisations
were generated using amount-based targets, amount-based and
surrogate-based targets, and amount-based and genetic-based targets for
each species. Data shows the performance of these prioritisations based
on how much genetic variation they explain. Bars denote means and
standard errors.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figures_files/figure-latex/unnamed-chunk-5-1.pdf}
\caption{The relationship between surrogates and genetic variation
secured in prioritizations.}
\end{figure}

>>>>

\setlength{\droptitle}{-2em}
  \title{Are environmental and geographic effective surrogates for genetic
variation in conservation planning?}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{Jeffrey O. Hanson$^1$, Jonathan R. Rhodes$^2$, Cynthia Riginos$^1$, Hugh
P. Possingham$^1$, Richard A. Fuller$^1$\\$^1$School of Biological
Sciences, The University of Queensland, Brisbane, QLD,
Australia\\$^2$School of Geography, Planning and Environmental
Management, The University of Queensland, Brisbane, QLD,
Australia\\Correspondance should be addressed to
\href{mailto:jeffrey.hanson@uqconnect.edu.au}{jeffrey.hanson@uqconnect.edu.au}}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{01 January 2016}

% load packages
\usepackage{amsmath,amsfonts,float,makecell,titletoc,tocloft,titlesec,natbib,pdfpages}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[utf8]{inputenc}

@@

First, we grouped individuals into genetic lineages for each species
(Figure S2). The purpose of this step was to avoid confounding variation
between populations with due to genetic history with variation
attributed to selection pressures. We used the Bayesian clustering
method implemented in $\texttt{Structure}$ (version 2.3.4; Pritchard
\emph{et al.} 2000; Falush \emph{et al.} 2007) assuming admixture and an
independent alleles model. For each species, we identified the most
plausible number of populations (K = 1--3; 4 runs per K; 10 burn-in
iterations; 10 post-burn-in iterations) using the $\Delta$K method
(Figure S3; Evanno \emph{et al.} 2005). We aggregated the population
membership probabilities from replicate runs using $\texttt{CLUMPP}$
(version 1.1.2; large K greedy search method with 1000 random starts per
K per species; Figure S4; Jakobsson and Rosenberg 2007). Individuals
with a moderate probability of belonging to a single population
($\geq 0.75$) were assigned accordingly (Figure S5), and subject to
subsequent outlier locus detection methods.

@@

\includepdf[pages=-]{figures.pdf}
\includepdf[pages=-]{supporting_information.pdf}

### 2016-01-01 article/supporting_information.Rmd
- set parameters to TOML format - make parameters loaded only in when needed to reduce overheads if parameters are changed

```{r, echo=FALSE}
plot.spp.mds <- function(i) {
	# define function
	make.mds.plot <- function(j,k) {
		ggplot() +
			geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
				fill='grey20', color='grey80') +
			geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat', 
				group='group', fill=paste0(unique(spp.samples.DF$species)[i], '_', j, '_d',k)),
				alpha=0.8, color='grey10') +
			guides(fill=guide_legend(title=' ')) +
			theme_classic() +
			theme(axis.ticks=element_blank(), axis.text=element_blank(),
				plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank()) +
			coord_cartesian(
				xlim=buffered.range(grid.FPLY$long, 0.05),
				ylim=buffered.range(grid.FPLY$lat, 0.05)
			) +
			xlab('') +
			ylab('') +
			ggtitle(paste0(j,' (',k,')'))
	}
	# init
	adapt.col <- paste0(unique(spp.samples.DF$species)[i], '_adaptive_d1')
	neutral.col <- paste0(unique(spp.samples.DF$species)[i], '_neutral_d1')
	plotLST <- list()
	# make adaptive plot
	if (adapt.col %in% names(grid.FPLY)) {
		for (k in seq_len(mds.k))
			plotLST <- append(plotLST, list(make.mds.plot('adaptive', k)))
	} 
	# make neutral plot
	if (neutral.col %in% names(grid.FPLY)) {
		for (k in seq_len(mds.k))
			plotLST <- append(plotLST, list(make.mds.plot('neutral', k)))
	} 
	# make plot
	do.call(
		grid.arrange,
		append(
			plotLST,
			list(ncol=6)
		)
	)
}
```

>>>>

```{r, echo=FALSE}
plot.spp.mds <- function(i) {
	# define function
	make.mds.plot <- function(j,k) {
		ggplot() +
			geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
				fill='grey20', color='grey80') +
			geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat', 
				group='group', fill=paste0(unique(spp.samples.DF$species)[i], '_', j, '_d',k)),
				alpha=0.8, color='grey10') +
			guides(fill=guide_legend(title=' ')) +
			theme_classic() +
			theme(axis.ticks=element_blank(), axis.text=element_blank(),
				plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank()) +
			coord_cartesian(
				xlim=buffered.range(grid.FPLY$long, 0.05),
				ylim=buffered.range(grid.FPLY$lat, 0.05)
			) +
			xlab('') +
			ylab('') +
			ggtitle(paste0(j,' (',k,')'))
	}
	# init
	adapt.col <- paste0(unique(spp.samples.DF$species)[i], '_adaptive_d1')
	neutral.col <- paste0(unique(spp.samples.DF$species)[i], '_neutral_d1')
	plotLST <- list()
	# make adaptive plot
	if (adapt.col %in% names(grid.FPLY)) {
		for (k in seq_len(mds.params.LST[[MODE]]$k))
			plotLST <- append(plotLST, list(make.mds.plot('adaptive', k)))
	} 
	# make neutral plot
	if (neutral.col %in% names(grid.FPLY)) {
		for (k in seq_len(mds.params.LST[[MODE]]$k))
			plotLST <- append(plotLST, list(make.mds.plot('neutral', k)))
	} 
	# make plot
	do.call(
		grid.arrange,
		append(
			plotLST,
			list(ncol=6)
		)
	)
}
```
